Learning by Teaching a Robot:\\The Case of Handwriting

### 2015-07-10 main.tex
Reworked Henry section

Thomas is five and an half, and has been diagnosed with visuo-constructive
deficits. He is under the care of an occupational therapist, and tries to
workaround his inability to draw letters in a consistent manner. Vincent is six
and struggles at school with his poor handwriting and even poorer
self-confidence.

@@

\begin{figure}[ht!]
    \centering
    \subfigure[Nine allographs of the cursive ``h'', next to the reference]{
        \includegraphics[width=0.4\columnwidth]{h}
    }
    \subfigure[Same samples, normalized and projected in the eigen space spanned from the
        3 first eigenvectors: clusters arise, that actually match writing styles.]{
        \includegraphics[width=0.4\columnwidth]{eigenspace-uniformization}
    }

@@

\subsubsection{Context, Study Design}

This study focused technical adaptations of the CoWriter activity for a
child diagnosed with real writing deficits.\\

Thomas is a six years old child. It is under the care of an occupational
therapist. According to the therapist, he shows visuo-constructive deficits.
As an effect in writing activities, he was frequently performing random attempts and then was comparing
with the provided template. What is more, Thomas is strongly careless : he rarely payed attention to
advices, even to what he was doing when he was currently drawing, and he was
quickly shifting his attention from one
activity to another.\\

Our objective is to investigate small modifications of the activity adapted to
the troubles of Thomas (visuo-constructive deficits and inattention) in order to
sustain him focused
on the activity during forty-minutes session, and to make the robot
evidently learning from his demonstrations.\\

The experiment took place in the therapist's office. It was divided in four
sessions. This time, we assumed that a scenario like the one we used for Vincent was
no longer relevant with Thomas. We just introduced the robot and quickly
said that it was seeking help to train for a robot handwriting contest.\\

Before the experiment, Thomas was working on writing numbers with the therapist.
Hence we decided to turn the CoWriter activity to teach numbers to the robot. Build on
existing iPad applications actually used by occupational therapists (Dawson Toth's \emph{ABC's
Writer}), we designed an Android application for pre-test and post-test. It consisted in drawing
numbers following an helping pattern, that becomes more fine with levels in
order to increase difficulty~\ref{fig:abc-writer}.\\

Since Thomas was frequently drawing horizontally-inverted numbers, or even
unrecognizable allographs, the learning algorithm of the robot was converging to
meaningless scrawls. To fix this problem, we programmed the robot to refuse allographs that
were too distant to a reference with a threshold we arbitrary fixed. In that way,
the child was forced to take care on what he was providing to the robot as
demonstration. Also, to make sure the robot was going to improve its handwriting
and to clearly show this improvement, we decided to make it starting
from scratch : for all numbers, the first try of the robot resulted in
a simple vertical stroke (see the first robot's try
in~\ref{learning_6_demos}).\\

In this setup, we added a second tablet with one button per numbers. It was used
by the child to chose a new number to teach to the robot. It also provided the
possiblity to enter letters or words, and to switch to another activity (the
robot telling a story).

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{abc-writer}
    \caption{\small Screenshot of the Android application developed to be used as
    pre-test and post-test: the child must follow with his finger the path of
the letter or number. We count the number of times the finger goes outside of
the red-bordered area.}
    \label{fig:abc-writer}
\end{figure}

\subsubsection{Results}

Despite his inattention, Thomas was able to keep on the activity during more than
forty minutes in each session. In total, 55 allographs out of 82 
provided by the child as demonstration were acceptable by the robot (with a
progressive improvement from 13 out of 28 in the first session up to 26 out
of 29 in the last session).\\

As soon as Thomas understood that the robot was only accepting recognizable
allographs, he started to take care on the allographs he was providing to it. He used the
clearing button to retry when he judged that his numbers where not going to be
accepted by the robot. According to the therapist, it was the first time Thomas
was intrinsically correcting himself.\\

Since the robot was learning from scratch, each time Thomas succeeded to make his
demonstrations accepted by it, the improvement of its handwriting was
visible~\ref{learning_6_demos},~\ref{henry_distances}.
Therefor, Thomas quickly entered in an self-rewarding game where the rewards were
the improvements of the robot, that enhanced his attention.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_demos}
    \caption{\small Thomas's demonstrations for the number ``6'' (top row) and
        corresponding shapes generated by the robot. After eight demonstrations,
        Thomas decided that the robot's ``6'' was good enough, and went to
    another character: in that respect, he was the one leading the learning
process of the robot.}
    \label{learning_6_demos}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_distances}
    \caption{\small Two metrics to assess the handwriting progresses: Euclidian
    distance in the eigenspace of the number dataset (top figure) or in
cartesian space (bottom figure). Green lines represent the robot performance,
blue lines Thomas's performance. The round IDs correspond to the demonstrations
pictured on Figure~\ref{henry_demos}.}
    \label{henry_distances}
\end{figure}

>>>>

Thomas is five and an half, and has been diagnosed with visuo-constructive
deficits. He is under the care of an occupational therapist, and tries to
workaround his inability to draw letters in a consistent manner. Vincent is six
and struggles at school with his poor handwriting and even poorer
self-confidence.\footnote{Children's names have been changed.}

@@

\begin{figure}[ht!]
    \centering
    \subfigure[Nine allographs of the cursive ``h'', next to the reference]{
        \includegraphics[width=0.4\columnwidth]{h}
    }
    \subfigure[Same samples, normalized and projected in the eigenspace spanned from the
        3 first eigenvectors: clusters arise, that actually match writing styles.]{
        \includegraphics[width=0.4\columnwidth]{eigenspace-uniformization}
    }

@@

The second long-term study was designed with the help of an occupational therapist in
Geneva and aimed at deploying the system for the first time on a real
therapeutic case.

Thomas is a 5.5 years old child, under the care of an occupational therapist. He
has been diagnosed with visuo-constructive deficits, which translate into
difficulties for him to consistently draw letters. Besides, focusing on a task
is difficult for Thomas, who tends to rapidly shift his attention to something
else. Since the robot's learning algorithm requires repeated demonstrations of
similarly shaped letters to converge, the occupational therapist was especially
interested in observing if the robot would induce a strong enough motivation for
Thomas to focus on producing regular, consistent letters, thus overcoming his
deficit.

The experiment took place at the therapist's surgery (four sessions spanning
over 5 weeks). Contrary to Vincent's experiment, we chose not to introduce any
storyline beyond a simple prompt (``the robot wants to participate to a robotic
handwriting contest, will you help him to prepare?''), only provided during the
first session. Hence, we also tested during this case-study if the robot (and
the Prot\'eg\'e effect) would induce a strong enough intrinsic motivation to
keep the child engage over the five weeks.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_demos}
    \caption{\small Thomas' demonstrations for the number ``6'' (top row) and
        the corresponding shapes generated by the robot (bottom row). After
        eight demonstrations, Thomas decided that the robot's ``6'' was good enough,
        and went to another character. In that regard, he was the one
        responsible for the learning process of the robot.}

    \label{learning_6_demos}
\end{figure}

The occupational therapist had recently carried activities with Thomas about
writing numbers, so we decided with her to focus on these as well: Thomas would
use a secondary tablet to tell the robot what number to write, and would then
correct the robot's attempts like in the other experiments.
Figure~\ref{learning_6_demos} shows the attempts/corrections cycles that
occurred during one of the session, on the number ``6''.

%Before the experiment, Thomas was working on writing numbers with the therapist.
%Hence we decided to turn the CoWriter activity to teach numbers to the robot.
%Build on existing iPad applications actually used by occupational therapists
%(Dawson Toth's \emph{ABC's Writer}), we designed an Android application for
%pre-test and post-test. It consisted in drawing numbers following an helping
%pattern, that becomes more fine with levels in order to increase
%difficulty~\ref{fig:abc-writer}.

Since Thomas would frequently draw mirrored numbers, or hard-to-recognize
shapes, the learning algorithm of the robot tended to converge towards
meaningless scribblings. We addressed this issue by having the robot to
\emph{refuse} allographs that were too far from the reference shape (\emph{``I'm
not sure to understand what you are drawing...''}), so that the child had to
pay good attention to what he would demonstrate to the robot. Also, to make
the robot's progresses evident, we modified the initialization step of the
learning algorithm to start with a roughly vertical stroke instead of a
deformed number (round 0 on Figure~\ref{learning_6_demos}).

%\begin{figure}
%    \centering
%    \includegraphics[width=0.9\linewidth]{abc-writer}
%    \caption{\small Screenshot of the Android application developed to be used as
%    pre-test and post-test: the child must follow with his finger the path of
%the letter or number. We count the number of times the finger goes outside of
%the red-bordered area.}
%    \label{fig:abc-writer}
%\end{figure}
%

Despite his attention deficit, Thomas was able to conduct the activity during
more than forty minutes in each session. In total, 55 allographs out of the 82
shapes demonstrated by the child were accepted by the robot (with a progressive
improvement from 13 out of 28 in the first session up to 26 out of 29 in the
last session).

As soon as Thomas understood that the robot was only accepting well-formed
allographs, he started to focus on it and he would typically draw 5 or 6 times
the number before actually sending to the robot (the tablet let the children
clear their drawing and try again before sending it to the robot). According to
the therapist, it was the first time that Thomas would correct himself in such a
way, explicitly having to reflect on how \emph{another agent} (the robot) would
interpret and understand his writing.

Since the robot's handwriting started from a simple primitive (a stroke), each
time Thomas succeeded to have his demonstration accepted by it, the robot's
improvement was clearly visible (as measured in Figure~\ref{henry_distances}).
This led to a self-rewarding situation that effectively supported Thomas'
engagement.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_distances}
    \caption{\small \textbf{Assessing handwriting progress}: Euclidian
    distance in the eigenspace of the number ``6''. Green lines represent the robot performance,
    blue lines Thomas's performance. The round IDs ($x$-axis) correspond to the demonstrations
pictured on Figure~\ref{learning_6_demos}.}
    \label{henry_distances}
\end{figure}
